State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-05-27 StateOfPractice/Papers/SOP_MI_OSR/SOP_MI_OSR.tex
Revision to SOP MI paper

We present the state of the practice for Medical Imaging (MI) software based on
data available in open source repositories. We selected 29 medical imaging
projects from 48 candidates and assessed 10 software qualities (installability,
correctness/ verifiability, reliability, robustness, usability, maintainability,
reusability, understandability, visibility/transparency and reproducibility) by
answering 108 questions for each software project. Based on the quantitative
data for the first 9 qualities, we ranked the MI software with the Analytic
Hierarchy Process (AHP). The four top ranked software products are \textit{3D
Slicer}, \textit{ImageJ}, \textit{Fiji}, and \textit{OHIF Viewer}.  Our ranking
is mostly consistent with the community's ranking, with four of our top five
projects also appearing in the top five of a list ordered by stars-per-year.  We
observed 88\% of the documentation artifacts recommended by research software
development guidelines.  However, the current state of the practice deviates
from the existing guidelines because of the rarity of some recommended artifacts
(like test plans, requirements specification, code of conduct, code style
guidelines, product roadmaps, and Application Program Interface (API)
documentation). Open source MI software projects may wish to consider increasing
their documentation.

@@

\noindent \emph{Keywords:}
	medical imaging, research software, software engineering, software
	quality, analytic hierarchy process, developer interviews

@@

We aim to study the state of software development practice for Medical Imaging
(MI) software using data available in open source repositories.  MI tools use
images of the interior of the body (from sources such as Magnetic Resonance
Imaging (MRI), Computed Tomography (CT), Positron Emission Tomography (PET) and
Ultrasound) to provide information for diagnostic, analytic, and medical
applications \citep{FDA2021, enwiki:1034887445, Zhang2008}.
Figure~\ref{Fig_Example}, which shows an image of the brain, highlights the
importance and value of MI. Through MI medical practitioners and researchers can
noninvasively gain insights into the human body, including information on
injuries and illnesses. Given the importance of MI software and the high number
of competing software projects, we wish to understand the merits and drawbacks
of the current development processes, tools, and methodologies. We aim to assess
through a software engineering lens the quality of the existing software with
the hope of highlighting standout examples, understanding current pain points
and providing guidelines and recommendations for future development.

@@

To gain insights, we devised 10 research questions, which can be applied to MI,
as well as to other domains, of research software \citep{SmithEtAl2021,
SmithAndMichalski2022}.  We designed the questions to learn about the
community's interest in, and experience with, software artifacts, tools,
principles, processes, methodologies, and qualities.  When we mention artifacts
we mean the documents, scripts and code that constitutes a software development
project. Example artifacts include requirements, specifications, user manuals,
unit tests, system tests, usability tests, build scripts, API (Application
Programming Interface) documentation, READMEs, license documents, process
documents, and code.  Once we have learned what MI developers do, we then put
this information in context by contrasting MI software against the trends shown
by developers in other research software communities.  Our aim is to collect
enough information to understand the current pain points experienced by the MI
software development community so that we can make some preliminary
recommendations for future improvements.

@@

\begin{enumerate}
	\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
	software projects exist, with the constraint that the source code must be
	available for all identified projects? (Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Which
	of the projects identified in \rqref{RQ_WhatProjects} follow current best
	practices, based on evidence found by experimenting with the software and
	searching the artifacts available in each project's repository?
	(Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
	similar is the list of top projects identified in \rqref{RQ_HighestQuality}
	to the most popular projects, as viewed by the scientific community?
	(Section~\ref{Sec_VsCommunityRanking})
    \item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
	do MI projects compare to research software in general with respect to the
	artifacts present in their repositories?
	(Section~\ref{Sec_CompareArtifacts})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareToolsProjMngmnt}:]
	How do MI projects compare to research software in general with respect to
	the use of tools (Section~\ref{Sec_CompareTools}) for:
	\begin{enumerate} 
		\item [\rqref{RQ_CompareToolsProjMngmnt}.a] development; and,
		\item [\rqref{RQ_CompareToolsProjMngmnt}.b] project management?
	\end{enumerate}
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareMethodologies}:]
	How do MI projects compare to research software in general with respect to
	principles, processes, and methodologies used?
	(Section~\ref{Sec_CompareMethodologies})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
	the pain points for developers working on MI software projects?
	(Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
	do the pain points of developers from MI compare to the pain points
	for research software in general? (Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For MI
	developers what specific best practices are taken to address the pain points
	and software quality concerns? (Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
	What research software development practice could potentially address the
	pain point concerns identified in \rqref{RQ_PainPoints}?
	(Section~\ref{ch_recommendations})

@@

With our methodology, we first choose a research software domain (in the current
case MI) and identify a list of about 30 software packages. (For measuring MI we
used 29 software packages.)  We approximately measure the qualities of each
package by filling in a grading template. Compared with our previous
methodology, the new methodology also includes repository based metrics, such as
the number of files, number of lines of code, percentage of issues that are
closed, etc.  With the quantitative data in the grading template, we rank the
software with the Analytic Hierarchy Process (AHP) (Section~\ref{ch_background}
provides details). After this, as another addition to our previous methodology,
we interview some development teams to further understand the status of their 
development process.

@@

To measure the existing MI software we need two sets of definitions: i) the
definitions of relevant software license models (Section
\ref{sec_software_categories}); and, ii) the definitions of the software
qualities that we will be assessing (Section \ref{sec_software_quality}). In our
assessment we rank the software packages for each quality; therefore, this
section also provides the background on our ranking process --- the Analytic
Hierarchy Process (Section \ref{sec_AHP}).

\subsection{Software Categories} \label{sec_software_categories}

When assessing software packages, we need to know the software's license.  In
particular, we need to know whether the source code will be available to us or
not.  We define three common software categories.  We will only assess software
that fits under the Open Source Software license.

\item \textbf{Open Source Software (OSS)} For OSS, the source code is openly
accessible. Users have the right to study, change and distribute it under a
license granted by the copyright holder. For many OSS projects, the development
process relies on the collaboration of different contributors worldwide
\citep{Corbly2014}. Accessible source code usually exposes more ``secrets'' of a
software project, such as the underlying logic of software functions, how
developers achieve their works, and the flaws and potential risks in the final
product. Thus, OSS is suitable for researchers analyzing the qualities of a
project.

\item \textbf{Freeware} Freeware is software that can be used free of charge.
Unlike OSS, the authors of do not allow access or modify the source code
\citep{LINFO2006}. To many end-users, the differences between freeware and OSS
may not be relevant. However, software developers who wish to modify the source
code, and researchers looking for insight into software development process may
find the inaccessible source code a problem.

\item \textbf{Commercial Software} ``Commercial software is software developed
by a business as part of its business'' \citep{GNU2019}. Typically speaking,
commercial software requires users to pay to access all of its features,
excluding access to the source code. However, some commercial software is also
free of charge \citep{GNU2019}. Based on our experience, most commercial
software products are not OSS.

@@

	\item \textbf{Reproducibility} ``A result is said to be reproducible if
	another researcher can take the original code and input data, execute it,
	and re-obtain the same result'' \citep{BenureauAndRougier2017}.
\end{itemize}

@@

We developed a methodology for evaluating the state of the practice of research
software \citep{SmithEtAl2021, SmithAndMichalski2022}.  The methodology can be
instantiated for a specific domain of scientific software, which in the current
case is medical imaging software for visualization.  Our methodology involves
and engages a domain expert partner throughout, as discussed in
Section~\ref{sec_vet_software_list}.  The four main steps of the methodology
are:

\begin{enumerate}
\item Identify list of representative software packages
(Section~\ref{sec_software_selection});
\item Measure (or grade) the selected software
(Section~\ref{sec_grading_software});
\item Interview developers (Section~\ref{sec_interview_methods});
\item Answer the research questions (as given in Section~\ref{sec_motivation}).
\end{enumerate}

@@

We grade the selected software using the measurement template summarized in
\citet{SmithEtAl2021}.  The template provides measures of the qualities listed
in Section~\ref{sec_software_quality}, except for reproducibility, which is
assessed through the developer interviews (Section~\ref{sec_interview_methods}).
For each software package, we fill in the template questions. To stay within the
target of 160 person hours to measure the domain, we allocated between one and
four hours for each package. Project developers can be contacted for help
regarding installation, if necessary, but we impose a cap of about two hours on
the installation process, to keep the overall measurement time feasible.
Figure~\ref{fg_grading_template_example} shows an excerpt of the spreadsheet.
The spreadsheet includes a column for each measured software package.

@@

\item \textbf{Correctness \& Verifiability} We check each project to identify
any techniques used to ensure this quality, such as literate programming,
automated testing, symbolic execution, model checking, unit tests, etc. We also
examine whether the projects use Continuous Integration and Continuous Delivery
(CI/CD). For verifiability, we go through the documents of the projects to check
for the presence of requirements specifications, theory manuals, and getting
started tutorials. If a getting started tutorial exists and provides expected
results, we follow it to check the correctness of the output.

@@

\subsection{Interview Methods} \label{sec_interview_methods}

The repository-based measurements summarize the information we can collect from
on-line resources. This information is incomplete because it doesn't generally
capture the development process, the developer pain points, the perceived
threats to software quality, and the developers' strategies to address these
threats.  Therefore, part of our methodology involves interviewing developers.

We based our interviews on a list of 20 questions, which can be found in
\citet{SmithEtAl2021}. Some questions are about the background of the software,
the development teams, the interviewees, and how they organize their projects.
We also ask about the developer's understanding of the users. Some questions
focus on the current and past difficulties, and the solutions the team has
found, or plan to try. We also discuss documentation, both with respect to how
it is currently done, and how it is perceived. A few questions are about
specific software qualities, such as maintainability, understandability,
usability, and reproducibility. The interviews are semi-structured based on the
question list; we ask follow-up questions when necessary.  The interview process
presented here was approved by the McMaster University Research Ethics Board
under the application number
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/MACREM/Application.pdf}
{MREB\#: 5219}.

We sent interview requests to all 29 projects using contact information from
projects websites, code repository, publications, and from biographic pages at
the teams' institutions.  In the end nine developers from eight of the projects
agreed to participate: \textit{3D Slicer}, \textit{INVESALIUS 3}, \textit{dwv},
\textit{BioImage Suite Web}, \textit{ITK-SNAP}, \textit{MRIcroGL},
\textit{Weasis}, and \textit{OHIF}. We spent about 90 minutes for each
interview. One participant was too busy to have an interview, so they wrote down
their answers. In one case two developers from the same project agreed to be
interviewed. We held the meetings on-line using either Zoom or Teams, which
facilitated recording and automatic transcription. The full interview answers
can be found in \citet{Dong2021}.

@@

@@

We analyzed the state of the practice for the MI domain with the goal of
understanding current practice, answering our ten research questions
(Section~\ref{sec_motivation}) and providing recommendations for current and
future projects.  Our methods in Section~\ref{ch_methods} form a general process
to evaluate domain-specific software, that we apply to the specific domain of MI
software. We identified 48 MI software candidates, then, with the help of the
Domain Expert selected 29 of them to our final list.

@@

For insight into devising future methods and tools, we interviewed nine
developers (from eight teams) to learn about their pain points
(Section~\ref{painpoints}).  We also discussed qualities of potential concern.
The identified pain points and qualities of concern include:

\begin{description}
\item [\ppref{P_LackDevTime}] Lack of development time, %P1
\item [\ppref{P_LackFunding}] Lack of funding, %P2
\item [\ppref{P_TechnologyHurdles}] Technology hurdles, %P3
\item [\ppref{P_Correctness}] Ensuring correctness, %P4
\item [\ppref{P_Usability}] Usability, %P5
\item [\qref{Q_Maintainability}] Quality of maintainability, and %Q1
\item [\qref{Q_Reproducibility}] Quality of reproducibility. %Q2
\end{description}

Despite the pain points, overall MI software is in a healthy state for software
development practices.  In our survey of the selected projects we observed 88\%
of the documentation artifacts recommended by research software development
guidelines (Section~\ref{Sec_CompareArtifacts}).  With respect to tools, MI is
keeping pace with other research software with 100\% of the projects using
version control (with 93\% specifically using git)
(Section~\ref{Sec_CompareTools}).  We observed that the MI developers tend to
follow the typical research software trend of using a quasi-agile software
development process (Section~\ref{Sec_CompareMethodologies}).

Although the state of the practice for MI software is healthy, we did notice
areas where practice seems to lag behind the research software development
guidelines.  For instance, the guidelines recommend three artifacts that were
not observed: uninstall instructions, test plans, and requirements
documentation. We observed the following recommended artifacts, but only rarely:
contributing file, developer code of conduct, code style guidelines, product
roadmap, design documentation, and API documentation
(Section~\ref{Sec_CompareArtifacts}). Although software development tool use
seems healthy, we found the use of CI/CD behind typical usage rates (17\% of the
projects used CI/CD) (Section~\ref{Sec_CompareTools}).  With respect to the
development process, developer identified areas for improvement included testing
(only 50\% of projects were identified to have unit testing) and documentation
(only three out of nine developers felt their documentation was clear enough)
(Section~\ref{Sec_CompareMethodologies}).

Our interviewees proposed strategies to improve the state of the practice, to
address the identified pain points, and to improve software quality.  To their
list (Section~\ref{painpoints}) we added some of our own recommended strategies
(Section~\ref{ch_recommendations}).  Below we summarize the proposed strategies,
with traceability to where we discuss the strategy, and to the relevant pain
points.

\item Increase documentation to address \ppref{P_LackDevTime},
\ppref{P_TechnologyHurdles}, \ppref{P_Usability}, \qref{Q_Maintainability},
\qref{Q_Reproducibility} (Section~\ref{painpoints})

\item Increase testing by enriching datasets to address \ppref{P_Correctness},
\qref{Q_Reproducibility}
(Section~\ref{painpoints},~\ref{sec_recommendations_testing_dataset})

\item Increase modularity to address \qref{Q_Maintainability} (Section~\ref{painpoints})

\item Use continuous integration to address \ppref{P_LackDevTime},
\ppref{P_Correctness}, \qref{Q_Reproducibility}
(Section~\ref{painpoints},~\ref{Sec_ContinuousIntegration})

\item Move to web applications to address \ppref{P_TechnologyHurdles}
(Section~\ref{painpoints},~\ref{sec_webapps})

\item Employ linters to address \ppref{P_LackDevTime},
\ppref{P_TechnologyHurdles}, \qref{Q_Maintainability}
(Section~\ref{Sec_Linters})

\item Peer reviews to address \ppref{P_LackDevTime},
\ppref{P_TechnologyHurdles}, \ppref{P_Correctness}, \qref{Q_Maintainability}
(Section~\ref{Sec_PeerReview})

\item Design for change to address \qref{Q_Maintainability} (Section~\ref{Sec_DesForChange})

\item Assurance case to address \ppref{P_Correctness}, \qref{Q_Reproducibility}
(Section~\ref{AssuranceCases})

\item Generate all things to address \ppref{P_LackDevTime},
\ppref{P_TechnologyHurdles}, \ppref{P_Correctness}, \ppref{P_Usability},
\qref{Q_Maintainability} and \qref{Q_Reproducibility}.
(Section~\ref{Sec_GenAllThings})

>>>>

We present the state of the practice for Medical Imaging (MI) software based on
data available in open source repositories. We selected 29 medical imaging
projects from 48 candidates and assessed 9 software qualities (installability,
correctness/ verifiability, reliability, robustness, usability, maintainability,
reusability, understandability, and visibility/transparency) by answering 108
questions for each software project. Based on the quantitative data, we ranked
the MI software with the Analytic Hierarchy Process (AHP). The top four software
products are \textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}, and
\textit{OHIF Viewer}.  Our ranking is mostly consistent with the community's
ranking, with four of our top five projects also appearing in the top five of a
list ordered by stars-per-year.  We observed 88\% of the documentation artifacts
recommended by research software development guidelines.  However, the current
state of the practice deviates from the existing guidelines because of the
rarity of some recommended artifacts (like test plans, requirements
specification, code of conduct, code style guidelines, product roadmaps, and
Application Program Interface (API) documentation).

@@

\noindent \emph{Keywords:}
	medical imaging, research software, software engineering, software
	quality, analytic hierarchy process

@@

We aim to study the state of software development practice for Medical Imaging
(MI) software using data available in open source repositories.  MI tools use
images of the interior of the body (from sources such as Magnetic Resonance
Imaging (MRI), Computed Tomography (CT), Positron Emission Tomography (PET) and
Ultrasound) to provide information for diagnostic, analytic, and medical
applications \citep{FDA2021, enwiki:1034887445, Zhang2008}.
Figure~\ref{Fig_Example}, which shows an image of the brain, highlights the
importance and value of MI. Through MI medical practitioners and researchers can
noninvasively gain insights into the human body, including information on
injuries and illnesses. Given the importance of MI software and the high number
of competing software projects, we wish to understand the merits and drawbacks
of the current development processes, tools, and methodologies. We aim to assess
through a software engineering lens the quality of the existing software with
the hope of highlighting standout examples, and providing guidelines and
recommendations for future development.

@@

To gain insight, we devised 10 research questions, which can be applied to MI,
as well as to other domains, of research software \citep{SmithEtAl2021,
SmithAndMichalski2022}.  We designed the questions to learn about the
community's interest in, and experience with, software artifacts, tools,
principles, processes, methodologies, and qualities.  When we mention artifacts
we mean the documents, scripts and code that constitutes a software development
project. Example artifacts include requirements, specifications, user manuals,
unit tests, system tests, usability tests, build scripts, API (Application
Programming Interface) documentation, READMEs, license documents, process
documents, and code.  Once we have learned what MI developers do, we then put
this information in context by contrasting MI software against the trends shown
by developers in other research software communities.

@@

\begin{enumerate}
	\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
	software projects exist, with the constraint that the source code must be
	available for all identified projects? (Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Which
	of the projects identified in \rqref{RQ_WhatProjects} follow current best
	practices, based on evidence found by experimenting with the software and
	searching the artifacts available in each project's repository?
	(Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
	similar is the list of top projects identified in \rqref{RQ_HighestQuality}
	to the most popular projects, as viewed by the scientific community?
	(Section~\ref{Sec_VsCommunityRanking})
    \item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
	do MI projects compare to research software in general with respect to the
	artifacts present in their repositories?
	(Section~\ref{Sec_CompareArtifacts})
\end{enumerate}

@@

With our methodology, we first choose a research software domain (in the current
case MI) and identify a list of about 30 software packages. (For measuring MI we
used 29 software packages.)  We approximately measure the qualities of each
package by filling in a grading template. Compared with our previous
methodology, the new methodology also includes repository based metrics, such as
the number of files, number of lines of code, percentage of issues that are
closed, etc.  With the quantitative data in the grading template, we rank the
software with the Analytic Hierarchy Process (AHP) (Section~\ref{ch_background}
provides details).

@@

To measure the existing MI software we need the definitions of the software
qualities that we will be assessing (Section \ref{sec_software_quality}). In our
assessment we rank the software packages for each quality; therefore, this
section also provides the background on our ranking process --- the Analytic
Hierarchy Process (Section \ref{sec_AHP}).

@@

@@

We developed a methodology for evaluating the state of the practice of research
software \citep{SmithEtAl2021, SmithAndMichalski2022}.  The methodology can be
instantiated for a specific domain of scientific software, which in the current
case is medical imaging software for visualization.  Our methodology involves
and engages a domain expert partner throughout, as discussed in
Section~\ref{sec_vet_software_list}.  The three main steps of the methodology
are:

\begin{enumerate}
\item Identify list of representative software packages
(Section~\ref{sec_software_selection});
\item Measure (or grade) the selected software
(Section~\ref{sec_grading_software});
\item Answer the research questions (as given in Section~\ref{sec_motivation}).
\end{enumerate}

@@

We grade the selected software using the measurement template summarized in
\citet{SmithEtAl2021}.  The template provides measures of the qualities listed
in Section~\ref{sec_software_quality}.
For each software package, we fill in the template questions. To stay within the
target of 160 person hours to measure the domain, we allocated between one and
four hours for each package. Project developers can be contacted for help
regarding installation, if necessary, but we impose a cap of about two hours on
the installation process, to keep the overall measurement time feasible.
Figure~\ref{fg_grading_template_example} shows an excerpt of the spreadsheet.
The spreadsheet includes a column for each measured software package.

@@

% move the definitions here; it seems repetitive the way it is now - maybe
% remove the definitions?
\item \textbf{Correctness \& Verifiability} We check each project to identify
any techniques used to ensure this quality, such as literate programming,
automated testing, symbolic execution, model checking, unit tests, etc. We also
examine whether the projects use Continuous Integration and Continuous Delivery
(CI/CD). For verifiability, we go through the documents of the projects to check
for the presence of requirements specifications, theory manuals, and getting
started tutorials. If a getting started tutorial exists and provides expected
results, we follow it to check the correctness of the output.

@@

@@

\section{Threats to Validity} \label{sec_threats_to_validity}

Below we categorize and list the threats to validity that we have identified.
Our categories come from an analysis of software engineering secondary studies
by \citet{AmpatzoglouEtAl2019}, where a secondary study analyzes the data from a
set of primary studies.  \citet{AmpatzoglouEtAl2019} is appropriate because a
common example of a secondary study is a systematic literature review. Our
methodology is a systematic software review --- the primary studies are the
software packages, and our work collects and analyzes these primary studies.  We
identified similar threats to validity in our assessment of the state of the
practice of Lattice Boltzmann Solvers \citep{SmithEtAl2024}.

\subsection{Reliability}

A study is reliable if repetition of the study by different researchers using
the original study's methodology would lead to the same results
\citep{RunesonAndHost2009}. Reliability means that data and analysis are
independent of the specific researcher(s) doing the study.  For the current
study the identified reliability related threats are as follows:

\begin{itemize}
\item One individual does the manual measures for all packages. A different
evaluator might find different results, due to differences in abilities,
experiences, and biases.
\item The manual measurements for the full set of packages took several months.
Over this time the software repositories may have changed and the reviewer's
judgement may have drifted.
\end{itemize}

In \citet{SmithEtAl2016} we reduced concern over the reliability risk associated
with the reviewer's judgement by demonstrating that the measurement process is
reasonably reproducible.  In \citet{SmithEtAl2016} we graded five software
products by two reviewers. Their rankings were almost identical. As long as each
grader uses consistent definitions, the relative comparisons in the AHP results
will be consistent between graders.

\subsection{Construct Validity}

\citet{RunesonAndHost2009} defines construct validity as the adopted
metrics representing what they are intended to measure. Our construct threats are
often related to how we assume our measurements influences the various software
qualities, as summarized in Section~\ref{sec_grading_software}. Specifically,
our construct validity related threats include the following:

\begin{itemize}
\item We make indirect measurement of software qualities since meaningful direct
measures for qualities like maintainability, reusability and verifiability, are
unavailable.  We follow the usual assumption that developers achieve higher
quality by following procedures and adhering to standards \citep[p.\
112]{VanVliet2000}.
\item As mentioned in Section~\ref{sec_result_installability}, we could not
install or build \textit{dwv}, \textit{GATE}, and \textit{DICOM Viewer}. We used
a deployed on-line version for \textit{dwv}, a VM version for \textit{GATE}, but
no alternative for \textit{DICOM Viewer}. We might underestimate their rank due
to these technical issues.
\item Measuring software robustness only involved two pieces of data. This is
likely part of the reason for limited variation in the robustness scores
(Figure~\ref{fg_robustness_scores}). We could add more robustness data by
pushing the software to deal with more unexpected situations, like a broken
Internet connection, but this would require a larger investment of measurement
time. 
\item We may have inaccurately estimated maintainability by assuming a higher
ratio of comments to source code improves maintainability. Moreover, we assumed
that maintainability is improved if a high percentage of issues are closed, but
a project may have a wealth of open issues, and still be maintainable.
\item We assess reusability by the number of code files and LOC per file. This
measure is indicative of modularity, but it does not necessarily mean a good
modularization. The modules may not be general enough to be easily reused, or
the formatting may be poor, or the understandability of the code may be low.
\item The understandability measure relies on 10 random source code files, but
the 10 files will not necessarily be representative. 
\item As discussed in Section~\ref{Sec_OverallQ}, our overall AHP ranking makes
the unrealistic assumption of equal weighting.
\item We approximated popularity by stars and watches
(Section~\ref{Sec_VsCommunityRanking}), but this assumption may not be valid. 
\item In building Table~\ref{Tbl_Guidelines} some judgement was necessary on our
part, since not all guidelines use the same names for artifacts that contain
essentially the same information.
\end{itemize}

\subsection{Internal Validity} \label{Sec_InternalValidity}

Internal validity means that discovered causal relations are trustworthy and
cannot be explained by other factors \citep{RunesonAndHost2009}. In our
methodology the internal validity threats include the following:

\begin{itemize}
\item In our search for software packages
(Section~\ref{sec_software_selection}), we may have missed a relevant package.
\item Our methodology assumes that all relevant software development activities
will leave a trace in the repositories, but this is not necessarily true. For
instance, the possibility exists that CI usage was higher than what we observed
through the artifacts (Section~\ref{Sec_CompareTools}). As another example,
although we saw little evidence of requirements
(Section~\ref{Sec_CompareTools}), maybe teams keep this kind of information
outside their repos, possibly in journal papers or technical reports.
\end{itemize}

\subsection{External Validity}

If the results of a study can be generalized (applied) to other
situations/cases, then the study is externally valid \citep{RunesonAndHost2009}.
We are confident that our search was exhaustive.  We do not believe that we
missed any highly popular examples.  Therefore, the bulk of our validity
concerns are internal (Section~\ref{Sec_InternalValidity}).
However, our hope is that the trends observed, and the lessons learned for MI
software can be applied to other research software.  With that in mind we
identified the following threat to external validity:

\begin{itemize}
\item We cannot generalize our results if the development of MI software is
fundamentally different from other research software.
\end{itemize}

Although there are differences, like the importance of data privacy for MI data,
we found the approach to developing LBM software \citep{SmithEtAl2024} and MI
software to be similar.  Except for the domain specific aspects, we believe that
the trends observed in the current study are externally valid for other research
software.

@@

We analyzed the state of the practice for the MI domain with the goal of
understanding current practice by answering our four research questions
(Section~\ref{sec_motivation}).  Our methods in Section~\ref{ch_methods} form a
general process to evaluate domain-specific software, that we apply to the
specific domain of MI software. We identified 48 MI software candidates, then,
with the help of the Domain Expert selected 29 of them to our final list.

@@

When it comes to recommended software artifacts, the state of open source MI projects is healthy, with our surveyed examples showing 88\%
of the documentation artifacts recommended by research software development
guidelines (Section~\ref{Sec_CompareArtifacts}).  However, we did notice
areas where practice seems to lag behind the research software development
guidelines.  For instance, the guidelines recommend three artifacts that were
not observed: uninstall instructions, test plans, and requirements
documentation. We observed the following recommended artifacts, but only rarely:
contributing file, developer code of conduct, code style guidelines, product
roadmap, design documentation, and API documentation
(Section~\ref{Sec_CompareArtifacts}).  Future MI projects may wish to consider incorporating more artifacts to potentially improve software quality.

Next steps.  Mention threats to validity?  Future work on pain points.  Point to
LBM methodology in that SOP paper?
